Using Visual Analytics to Explain\\[1ex] Black-Box Machine Learning

### 2018-03-05 tex/introduction/machine_learning.tex
health care and visual analytics

\subsection{Failure Modes}
\todo{failures}
- overfitting
- underfitting
- data
- model

\section{Visual Analytics}
% Visual analytics is the area of studying interactive graphical representations of abstract information with the help of computational and statistical procedures, in order to effectively analyze complex data. With regards to understanding the decision making of predictive machine learning models, there are two main strategies: white-box and black-box analysis \cite{class_signatures}.
\todo{visual analytics}
- use of high processing power of human eye
- interaction
- principles like Gestalt principles or preattentive perception or cleveland mcgill
% http://info.slis.indiana.edu/~katy/S637-S11/cleveland84.pdf
- ancombes quartett

>>>>

\subsection{Failure Modes}
\todo{failures}
- overfitting
- underfitting
- data
- model
- unexpected and different from how humans would fail

\section{Visual Analytics}
Gaining meaningful insights about large datasets is hard.
While computing various statistics about the data is often helpful it might oversimplify or in some cases even mislead information about the data.
One example dataset often used to demonstrate this is Ancombe's quartett \todo{cite ancombe}.
The quartett consists of four different datasets that all have the same mean, standard deviation, \todo{what else?}.
However, plotting the values of the datasets reveals that each one of those datasets obeys a different, unique characteriztic \todo{image}.

Visual analytics is the area of studying interactive graphical respresentations for analyzing complex data such as Ancombe's quartett mentioned above.
This is often done in addition to and with the help of computational and statistical procedures.
Various studies (\todo{cite}
% http://info.slis.indiana.edu/~katy/S637-S11/cleveland84.pdf
% preattentive, gestalt
) have been performed to establish visual principles to effectively take advantage of the high processing power of the human visual cortex \todo{?}.
Additionally, interaction with the graphical representations allows to hide some of the information and reveal it through interaction to not overwhelm the analyst.
For example, ``overview first; details on demand" \todo{shneiderman} is a popular mantra taking advantage of interaction.

In the context of machine learning, visual analytics, at first, seems to be a roadblock.
Integrating a human in a machine learning workflow inevitable leads to said human being the bottleneck.
Humans are magnitudes slower than computers which have to idle while waiting for human input.
Thus, human input should only be required for tasks that are impossible without.

\todo{maybe mention hyperparameter / architecture here}
For example, with the goal of improving the accuracy of a model a human should not have the task of, \eg, manually picking the order of the features to be included in a decision tree.
A good intuition might outperform the computer momentarily but having an objective goal makes it possible to eventually find a suitable algorithm that is en par or even outperforms the human.
Additionally, manual decisions are not generalizable to other tasks and not scalable to larger datasets which in turn is the reason to use machine learning to begin with.

However, \todo{}

- accuracy vs. semantic accuracy
- debugging models / data
- domain expert
- understanding /trusting models

### 2018-03-05 tex/introduction/usecase.tex
health care and visual analytics

\section{Use Case: Machine Learning for Health Care}
\todo{health}
* inconsistent coding NDC, IDC, hierarchies
* Doctor decision tree

>>>>

\section{Use Case: Machine Learning for Health Care}
In this thesis we will mostly explore use cases from the medical domain\footnote{However, results are easily transferable to other domains as well.}.
It is quite interesting to see parallels between the decision making process of doctors and machine learning.

Over thousands of years the health care domain has evolved into what we now know as evidence based medicine.
That is, new drugs can only be approved after long studies that show significant improvements and methods have been developed and refined over time to improve their measurable effectiveness.

For example, when a patient is examined by a doctor, the doctor follows a structured approach similar to a decision tree.
The first few nodes in the tree set up the context: age, ethnicity, weight, visual appearance, \etc.
A young patient has a vastly different set of potential health risks than an old patient.
Likewise, a physically active patient has different health risks than an obese patient.
Only after those initial features come symptoms into play.
This pattern is also reflected in the documentation of a patient encounter, the doctor's note.
Doctor's notes are structured in a way that reinforces this decision tree approach by putting the context establishing features first.

However, when it comes \todo{}

* Doctor decision tree (doctor note, limitations, residents vs. attending intuition; seniors might not be able to fully explain decisions made because of learned examples)
* inconsistent coding NDC, IDC, hierarchies
